---
title: "Open Policing Project Tutorial in Wichita, Kansas"
author: "Zandt Durham"
output:
  html_document:
    code_folding: hide
---

```{r setup, message=FALSE,  warning=FALSE}
## Libraries to include
library(tidyverse)
library(lubridate)

## Load the data
# Replace the path below with the path to where your data lives
data_path <- "https://datajournalism.tech/wp-content/uploads/2019/10/wichita.csv"
stops <- read_csv(data_path)

# Additional data and fixed values we'll be using
population_2016 <- tibble(
  subject_race = c(
    "asian/pacific islander", "black", "hispanic", "other/unknown","white"
  ),
  num_people = c(19294, 42485, 65090, 16686, 245499)
) %>% 
  mutate(subject_race = as.factor(subject_race))

center_lat <- 37.689789
center_lng <- -97.322720
```

## Exploratory data analysis


```{r}
colnames(stops)
nrow(stops)
str(stops)
summary(stops)
```

```{r}
stops %>% 
  group_by(subject_race) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
```



## Benchmark test

We saw before that over two-thirds of stops were of black drivers. The by-race
stop counts are only meaningful, though, when compared to some baseline. If
the Philadelphia population was about two-thirds black, then two-thirds of stops
being of black drivers wouldn't be at all surprising. 

```{r}
population_2016 %>% 
  mutate(prop = num_people / sum(num_people))
```


```{r}
stops %>% 
  count(subject_race) %>% 
  left_join(
    population_2016,
    by = "subject_race"
  ) %>% 
  mutate(stop_rate = n / num_people)
```


## Data Visualization

###Bar Chart
```{r bar-chart}
#data prep

barprep <- stops %>% 
  group_by(subject_race) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
#make the plot

bar <- ggplot(barprep,
       aes(x=reorder(subject_race,n), y=n))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="green")+
  geom_hline(yintercept = 0) +
  labs(title="Stopped Drivers by Race",
       subtitle = "Caucasian drivers were stopped the most in Wichita, Kansas")+
  coord_flip()

options(scipen=10000)

bar
```

###Bar Chart
```{r bar-chart-3}
#data prep

barprep3 <- stops %>% 
  group_by(subject_sex) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
#make the plot

bar3 <- ggplot(barprep3,
       aes(x=reorder(subject_sex,n), y=n))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="pink")+
  geom_hline(yintercept = 0) +
  labs(title="Stopped Drivers by Gender",
       subtitle = "Male drivers were stopped the most Wichita, Kansas")+
  coord_flip()

options(scipen=10000)

bar3
```

###Bar Chart
```{r bar-chart-2}
#data prep

barprep2 <- stops %>% 
  group_by(raw_defendant_ethnicity) %>% 
  summarize(
    n = n(),
    prop = n / nrow(.)
  )
#make the plot

bar2 <- ggplot(barprep2,
       aes(x=reorder(raw_defendant_ethnicity,n), y=n))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="blue")+
  geom_hline(yintercept = 0) +
  labs(title="Raw_defendant_ethnicity",
       subtitle = " 'N' drivers were stopped the most in Wichita, Kansas")+
  coord_flip()

options(scipen=10000)

bar2
```